---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Minh Vu"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
install.packages("readr")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("skimr")


library(readr)
library(dplyr)

# Reading data
cookies <- read_csv("choc_chip_cookie_ingredients.csv",
                    col_types = cols(
                      Ingredient = col_character(),
                      Text = col_character(),
                      Recipe_Index = col_character(),
                      Rating = col_double(),
                      Quantity = col_double(),
                      Unit = col_character()
                    ))

glimpse(cookies)

library(reticulate)

# Install pip into the reticulate environment
py_install("pip")

# Then install pandas
py_install("pandas")


```

```{python}
import sys
!{sys.executable} -m pip install pandas

import pandas as pd

cookies = pd.read_csv(
    "choc_chip_cookie_ingredients.csv",
    dtype={
        "Ingredient": "string",
        "Text": "string",
        "Recipe_Index": "string",
        "Unit": "string"
    }
)

cookies["Rating"] = pd.to_numeric(cookies["Rating"], errors="coerce")
cookies["Quantity"] = pd.to_numeric(cookies["Quantity"], errors="coerce")

cookies.info()
cookies.head()


```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1.What ingredients appear in the highest proportion of recipes?

2.Are higher-rated recipes associated with specific ingredients?

3.How much variation exists in the quantity of key ingredients like sugar or butter?

4.Are there ingredients that strongly co-occur together?

5.Do recipes from different sources (AllRecipes vs Epicurious vs Misc) differ in ingredient usage?


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
install.packages("skimr")  # if needed
library(skimr)

skim(cookies, 
     numeric.min = 0,
     histogram = TRUE)

```

```{python}
 pip install skimpy  

from skimpy import skim

skim(cookies)

```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.
```{r}
total_recipes <- n_distinct(cookies$Recipe_Index)

ingredient_props <- cookies %>%
  distinct(Recipe_Index, Ingredient) %>%
  count(Ingredient, sort = TRUE) %>%
  mutate(proportion = n / total_recipes) %>%
  slice_max(n, n = 20)

ingredient_props
```

```{python}
total_recipes = cookies["Recipe_Index"].nunique()

ingredient_props = (
    cookies[["Recipe_Index", "Ingredient"]]
    .drop_duplicates()
    .groupby("Ingredient")
    .size()
    .reset_index(name="count")
)

ingredient_props["proportion"] = ingredient_props["count"] / total_recipes

top20 = ingredient_props.sort_values("count", ascending=False).head(20)

top20

```



[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.
```{r}
rare_ingredients <- cookies %>%
  distinct(Recipe_Index, Ingredient) %>%
  count(Ingredient) %>%
  filter(n < 20)

paste(rare_ingredients$Ingredient, collapse = ", ")

```
```{python}
rare = ingredient_props[ingredient_props["count"] < 20]["Ingredient"]

print(", ".join(rare))

```


(Delete this note, but you can include data values inline in markdown text by using backticks, at least in R. For instance, here is R's built in value for pi: `r pi`. Unfortunately, this doesn't work in python using the knitr markdown engine, but you can print the list out in python anyways using a code chunk.)

### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.
```{r}
library(ggplot2)

ggplot(ingredient_props, aes(x = reorder(Ingredient, proportion), 
                             y = proportion)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 20 Ingredients by Proportion of Recipes",
       x = "Ingredient",
       y = "Proportion of Recipes")
```
```{python}
import matplotlib.pyplot as plt

top20_sorted = top20.sort_values("proportion")

plt.figure()
plt.barh(top20_sorted["Ingredient"], top20_sorted["proportion"])
plt.xlabel("Proportion of Recipes")
plt.title("Top 20 Ingredients by Proportion of Recipes")
plt.show()

```

[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

From this plot, we can see that ingredients like all-purpose flour, butter, and chocolate chips appear in nearly all recipes, as expected. However, ingredients like brown sugar and vanilla extract show slightly more variation, which may reflect differences in cookie style or regional preferences. A next step could be analyzing how ingredient presence relates to recipe rating.

